---
title: "Homework 2"
author: "Jinghao Zhang"
date: "2026-10-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

## Problem 1

```{r}
nyc_transit =
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(
    entry = recode_values(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
```

This dataset includes line, station name, latitude and longitude, routes, entry, vending, entrance type, and ADA. I selected the variables needed for the analysis, and converted `entry` from YES/NO to TRUE/FALSE. The cleaned dataset has 1868 rows and 19 columns. The data are not tidy because the routes are stored in multiple columns.

```{r}
nyc_transit |>
  distinct(line, station_name)
```


```{r}
nyc_transit |>
  filter(ada == TRUE) |>
  distinct(line, station_name)
```


```{r}
nyc_transit |>
  filter(vending == "NO") |>
  filter(entry == TRUE)
```

proportion of station entrances / exits without vending allow entrance = 69/183 = 37.7%

```{r}
nyc_transit_tidy = 
  nyc_transit |>
  mutate(
    across(route1:route11, as.character)) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name")

nyc_transit_tidy
```

```{r}
nyc_transit_tidy |>
  filter(route_name == "A") |>
  distinct(line, station_name)
```

there are 60 distinct stations.

```{r}
nyc_transit_tidy |>
  filter(route_name == "A", ada == TRUE) |>
  distinct(line, station_name)
```

17 stations are ADA compliant.

