---
title: "Homework 2"
author: "Jinghao Zhang"
date: "2026-10-07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```

## Problem 1

```{r}
nyc_transit =
  read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") |>
  janitor::clean_names() |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(
    entry = recode_values(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )
```

This dataset includes line, station name, latitude and longitude, routes, entry, vending, entrance type, and ADA. I selected the variables needed for the analysis, and converted `entry` from YES/NO to TRUE/FALSE. The cleaned dataset has 1868 rows and 19 columns. The data are not tidy because the routes are stored in multiple columns.

```{r}
nyc_transit |>
  distinct(line, station_name)
```


```{r}
nyc_transit |>
  filter(ada == TRUE) |>
  distinct(line, station_name)
```


```{r}
nyc_transit |>
  filter(vending == "NO") |>
  filter(entry == TRUE)
```

proportion of station entrances / exits without vending allow entrance = 69/183 = 37.7%

```{r}
nyc_transit_tidy = 
  nyc_transit |>
  mutate(
    across(route1:route11, as.character)) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name")

nyc_transit_tidy
```

```{r}
nyc_transit_tidy |>
  filter(route_name == "A") |>
  distinct(line, station_name)
```

there are 60 distinct stations.

```{r}
nyc_transit_tidy |>
  filter(route_name == "A", ada == TRUE) |>
  distinct(line, station_name)
```

17 stations are ADA compliant.

## Problem 2
```{r}
mr_trash_wheel =
  read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.integer(year),
    sports_balls = as.integer(round(sports_balls)),
    trash_wheel = "Mr. Trash Wheel"
  )

professor_trash_wheel =
  read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = cell_cols("A:M"),
    na = "dive"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.integer(year),
    trash_wheel = "Professor Trash Wheel"
  )

gwynnda =
  read_excel(
    "data/202610 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynnda the Good Wheel of the W",
    range = cell_cols("A:L")
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    year = as.integer(year),
    trash_wheel = "Gwynnda"
  )

trash_wheel_tidy =
  bind_rows(
    mr_trash_wheel,
    professor_trash_wheel,
    gwynnda
  )

trash_wheel_tidy
```

```{r}
professor_total_weight =
  trash_wheel_tidy |>
  filter(trash_wheel == "Professor Trash Wheel") |>
  summarize(total_weight = sum(weight_tons, na.rm = TRUE))

gwynnda_cigarette_butts =
  trash_wheel_tidy |>
  filter(
    trash_wheel == "Gwynnda",
    year == 2022,
    month == "June"
  ) |>
  summarize(total_butts = sum(cigarette_butts, na.rm = TRUE))

professor_total_weight

gwynnda_cigarette_butts
```
The combined dataset contains 1,269 observations and 15 variables, including key varibles like dumpster, month, year, and weight. Professor Trash Wheel collected 295 tons of trash. 18,120 cigarette butts were collected by Gwynnda in June 2022.
